Title Goes Here

### 2021-04-09 analysis/paper/shceof.Rmd
Reemplaza gráfico PSA vs EOF2 por tabla de correlación
#34

(ref:psa-eof2-cap) Relationship between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies [@mo2001]. For the period 1979 -- 2019.

@@

cEOF2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  # .[, hgt := rotate(hgt, psa_angle)] %>%
  sep_ReIm() %>% 
  ggplot(aes(value, hgt)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_smooth(method = "lm") + 
  scale_y_continuous("EOF2") +
  scale_x_continuous("PSA") +
  facet_grid(part ~ PC) +
  panel_background +
  tag_facets("rc")
```

The relationship between PSAs and the modes described from the cEOFs is also studied. Figure\ \@ref(fig:psa-eof2) shows the relationship between the two PSA indices and the Real and Imaginary phase of cEOF2. As anticipated by Figure\ \@ref(fig:eof2-regr-gh), there is a strong correlation between PSA1 and Imaginary cEOF2 (Figure\ \@ref(fig:psa-eof2)b.1), and between PSA2 and Real cEOF2 (Figure\ \@ref(fig:psa-eof2)a.2). Conversely, there is no relationship between PSA1 and Real cEOF2, and between PSA2 and Imaginary cEOF2 (Figure\ \@ref(fig:psa-eof2) panels a.1 and b.2). So not only this cEOF2 represents well both the spatial structure and temporal evolution of the PSA modes, but it's also possible to make a rather clean association between its two phases and the two PSA modes.

>>>>

(ref:psa-eof2-cap) Correlation between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies [@mo2001]. For the period 1979 -- 2019.

@@

make_numeric <- function(x) {
  x <- strsplit(x, "\ ")
  vapply(x, function(x) as.numeric(x[[1]]), numeric(1))
}

background_palette <- grDevices::colorRampPalette(c(scales::muted("blue"),
                                                    "white",
                                                    scales::muted("red")),
                                                  space = "Lab")

cEOF2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  sep_ReIm() %>% 
  .[, correlate(value, hgt), by = .(PC, part)] %>% 
  dcast(PC ~ part, value.var = "text") %>% 
  kbl2(booktabs = TRUE, caption = "(ref:psa-eof2-cap)") %>%
  column_spec2(2:3, background = spec_color2(make_numeric(.table[[.col]]),
                                             option = background_palette,
                                             scale_from = c(-1, 1)), 
               color = ifelse(abs(make_numeric(.table[[.col]])) > .5, "white", "black")) %>%
  add_header_above(c("", "cEOF2" = 2)) %>%
  kable_classic()

# cEOF2[psa_time, on = "time"] %>% 
#   .[PC != "SAM"] %>% 
#   copy() %>% 
#   # .[, hgt := rotate(hgt, psa_angle)] %>%
#   sep_ReIm() %>% 
#   ggplot(aes(value, hgt)) +
#   geom_hline(yintercept = 0, size = 0.2, color = "gray50") +
#   geom_vline(xintercept = 0, size = 0.2, color = "gray50") +
#   geom_point() +
#   geom_smooth(method = "lm") +
#   geom_smooth(method = "lm") + 
#   scale_y_continuous("EOF2") +
#   scale_x_continuous("PSA") +
#   facet_grid(part ~ PC) +
#   panel_background +
#   tag_facets("rc")
```

The relationship between PSAs and the modes described from the cEOFs is also studied. Figure\ \@ref(tab:psa-eof2) shows the relationship between the two PSA indices and the Real and Imaginary phase of cEOF2. As anticipated by Figure\ \@ref(fig:eof2-regr-gh), there is a strong correlation between PSA1 and Imaginary cEOF2, and between PSA2 and Real cEOF2. Conversely, there is no relationship between PSA1 and Real cEOF2, and between PSA2 and Imaginary cEOF2. So not only this cEOF2 represents well both the spatial structure and temporal evolution of the PSA modes, but it's also possible to make a rather clean association between its two phases and the two PSA modes.
